Scaffolder - Software for Microbial Genome Scaffolding.

### 2011-01-26 text/src/article.tex
Revision of first half of implementation section

% }}}
% {{{ Background
\section*{Background}
 Text for this section.\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}
% }}}
% {{{ Implementation
%%%%%%%%%%%%%%%%%%%%%%
\section*{Implementation}
  Text for this section \ldots
% }}}
\section*{Results and Discussion} %{{{1

\subsection*{Scaffolder Description} %{{{2

Individual reads produced from nucleotide sequencing can be assembled into
contiguous regions (``contigs'') of many overlapping individual reads. These
reads may however not represent a complete genome sequence where all contigs do
not result in a single complete sequence. The software outlined here
``Scaffolder'' allows these contigs to be further joined together into
a continuous single sequence. This selection, order and direction of contigs
into a final draft genome is often described as the `scaffold'. Scaffolder
allows the manual or computational specification of a genome scaffold by
specifying which contig sequences are used in simple to edit configuration
files.

The scaffold file specifies how contigs should be joined together in the draft
genome sequence. The raw nucleotide sequence for each contig is stored in
a fasta file, as might be produced from assembly of sequencing reads using
programs such as newbler\cite{newbler} or abyss\cite{abyss}. Scaffolder then
the contig to be selected by specifying the name of the contig in the scaffold
file. The scaffold file also allows a sub-sequence or the reverse complement of
a contig to used. This therefore preserves the original sequence data but also
allowing sequences to manipulated in the final genome build.

The majority of a genome can be estimated from scaffolding assembled contigs
into the most parsimonious build. There may however still be internal gaps in
the contigs from mate pair assembly based on estimated mate pair insert size.
Scaffolder allows inserts to be added to scaffolded so that gaps may be bridged
and filled - a common process in finishing a genome sequence. Inserts may be
treated the same way as a contig where the raw sequence, as may be produced by
PCR is stored in a fasta file, and then trimmed and/or reversed to fill the gap
in the contig.

Scaffolder allows the specification of unknown regions in the scaffold. These
are areas of the genome where the nucleotide sequence is unknown but the
approximate size region may be estimated from mate pair insert size or from
mapping contigs to reference genome and estimating the size of the gap. These
unknown regions allow a continuous draft sequence to be scaffolded but whilst
still highlighting regions waiting to be resolved.

\subsection*{Defining a scaffold in a text file} %{{{2

The genome scaffold is defined YAML format. YAML is an easy to edit data format
that is both machine readable and easy to edit manually.

INSERT FIGURE WITH EXAMPLE LAYOUT OF SCAFFOLD

  * Screen shot with example file
  * Illustrate three types of region in the scaffold file
  * Explanation of what each region does
  * Expected output from each region

Scaffolder provides a ruby API that allows the scaffold and designated
sequences to be manipulated through a ruby programme. This allows scaffolder to
be integrated into exiting genomics works flows such as using Rake. Scaffolder
all provides a command line binary for the manipulation of the scaffold. This
binary behaves as a standard Unix command line interface returning appropriate
exit statuses and providing manual pages for each of the commands. These two
methods are outlined in the following two sections.

\subsection*{Command line usage} %{{{2

  * Illustrate running command line to generate output
  * Demonstrate scaffold validate to test inserts overlap

\subsection*{Programmatic usage of API} %{{{2

  * Demonstrate how scaffolder can be manipulated programmatically
  * Examine scaffolder data using a ruby program
  * Example of extending sequence class with a new method \%GC content
  * Create a new class to do something

\subsection*{Installation} %{{{2

  * Install through RubyGems
  * Make sure the bin is in the class path

\section*{Conclusions} %{{{1
  Text for this section \ldots
% }}}
% {{{ Availability and Requirements
\section*{Availability and Requirements}

>>>>

% }}}
% {{{ Background
\section*{Background}
 Text for this section.
% }}}
\section*{Implementation} %{{{1

\subsection*{Code and Dependencies} %{{{2

Scaffolder is written in the Ruby programming language, version 1.8.7
\cite{ruby-lang}. Unit tests were implemented during development to maintain
functionality of individual elements of source code. Integration tests were
also written to test the scaffold software interface as a whole. These unit and
integration tests were written using the RSpec and Cucumber libraries
respectively\cite{rspec}.\pb

The Scaffolder source code is documented using the Yard library\cite{yard}.
Unix manual pages for the command line interface were generated using the Ronn
library\cite{ronn}. The manipulation of biological sequences in Scaffolder is
performed using the BioRuby library\cite{bioruby}. A full list of the software
dependencies in scaffolder can be found in the Gemfile in the root of source
code directory.\pb

\subsection*{Scaffold file syntax} %{{{2

The specification of the nucleotide sequences comprising the genome scaffold is
written using YAML \cite{yaml}. YAML is a data format using whitespace and
identation to produce a standardised machine readable structure. As a commonly
accepted data format third-party developers have the option to generate
a genome scaffold using their programming language of choice. As a standardised
format YAML formatted scaffold files can checked for syntax validity using
third party tools such as Kwalify\cite{kwalify}. \pb

In addition to being machine readable it also important that a genome scaffold
file also be easy to edit manually. As such this was the second reason the YAML
syntax was selected for the scaffold file data format. YAML has a concise and
compact syntax and when compared with formats such as extensible markup
language (XML) which require angular braces (`$<$' and `$>$') and open and
closing tags to delimite each data field.

By choosing a human readable syntax it is aimed that the scaffold should be
easier to edit and change manually in simple text edit.  A scenario which we
anticipate is common in the on-going process of scaffolding a genome sequence.
\pb

The entries in the YAML scaffold are defined as a list. Each entry in the list
corresponds to a region sequence in the final output sequence which follows the
previous entry in the scaffold and precedes the following entry.  Each
attribute defined for each entry in the list allows specifications for the
entry such as if any starting sequence should be clipped from the beginning or
end, or if the sequence should be reversed. \pb

INSERT IMAGE OF SCAFFOLD GRAPH AND RESULTING OUTPUT.

\subsection*{Parsing the YAML scaffold} %{{{2

Each stretch of sequence in the genome scaffold is defined as a list entry in
the YAML scaffold file. When parsing the scaffold file Scaffolder iterates over
each scaffold entry to create an instance of the Scaffolder::Region class.\pb

The Region class is inherited by all types of sequence in the scaffold. The
Region provides helper methods common to the tasks in manipulating scaffold
sequence, subsequencing and reversing. In particular the region class has
a class annotation method\cite{class-annotations} which allows attributes in
the scaffold method to be defined in the class. This negates having to write
a custom method to deal with each scaffold variable.  Furthermore this makes
the process of transferring attributes from the scaffold file to each region
instance much simpler - each yaml attribute corresponds to a method call on the
instantiated class. The scaffold file effectively becomes a DSL for populating
the scaffold region instances while still maintaining the machine readable data
qualities of YAML.\pb

The region class provides a sequence hook method. This method returns by
default the sequence corresponding to the fasta sequence defined by the source
tag which is then sublclassed or reversed if necessary. Subclasses of the
region class should override this method to provide custom functionality in
manipulating the source sequence. For example, the Scaffolder::Region::Sequence
class overrides the sequence method add any sequence inserts specified in the
scaffold file.\pb

\section*{Results and Discussion} %{{{1
  \subsection*{Results sub-heading}
    \subsubsection*{This is a sub-sub-heading}
      Sub-sub-sub-headings are made with the \textsl{\\subsubsection} command. \pb
      pb at end of lines ensures correct paragraph spacing.\pb
	  Text for this sub-sub-section \ldots
    \subsubsection*{Another sub-sub-sub-heading}
      Text for this sub-sub-section \ldots

  \subsection*{Another results sub-heading}
    Text for this sub-section \ldots

  \subsection*{Yet another results sub-heading}
    Text for this sub-section.  More results \ldots
% }}}
% {{{ Conclusions
%%%%%%%%%%%%%%%%%%%%%%
\section*{Conclusions}
  Text for this section \ldots
% }}}
% {{{ Availability and Requirements
\section*{Availability and Requirements}
